%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape
    If I have seen further it is by standing on the shoulders of Giants.} \\ \medskip
    --- Isaac Newton
\end{flushright}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
Firstly, I would like to thank my supervisors Prof. Steven Niederer and Dr. Anna Sher, for helping me become the scientist I am today. Thank you, Steve, for showing me the way to do good research, for being humble and always straight to the point. Thank you, Anna, for always wanting me to aim for perfection, for caring about every little detail and for your precious advice on the world of pharma. You both have been of inspiration to me, and I will cherish the knowledge you have transmitted to me for years to come.

\vspace{0.2cm}
I would like to express my gratitude to my examiners Prof. Peter Challenor and Prof. Gary Miriams, for dedicating their time to improving this research's quality through their valuable expertise.

\vspace{0.2cm}
I would like to thank all my Cardiac Electromechanics Research Group colleagues, with whom I shared such good times and built a solid friendship. In particular, thank you, Marina, for having been in the right place at the right moment. Without you, I would have never started a PhD in London, and none of this would have ever happened. Thank you, Cristobal, for being my first Spanish friend and sharing almost all our PhD "adventures'' in and out of the office. Thank you, Alex, for helping me through the initial tough times of my PhD when I was lost and for your patience and wise pieces of advice. Thank you, Hugh, for improving my English communication skills by only talking to you -- jokes apart -- for your resourcefulness and sharp-wittedness and your espresso martinis. Thank you, Adelisa, for bringing joy to the office on rainy days (basically every day), for your incredible stories and your open-mindedness. Thank you, Damiano, for being my Italian buddy with whom I never missed a coffee at lunchtime or a beer at night. Thank you, Sofia, for your stubbornness, loyalty and kindness.

\vspace{0.2cm}
Lastly, I would like to thank my family for always believing in me and for their continuous support throughout all these years far from home, and to Regina, for supporting all my life choices no matter how difficult it could have been.

\endgroup